---
title: "Danish Kings"
author: "Mathilde Petersen, Josefine Topbjerg, Freja Ipsen, Sissel Breckling"
date: "2026-05-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the kings

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator: ,  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.

```{r eval=TRUE}
# FILL IN THE CODE BELOW
library(tidyverse)
kings1 <- read.csv("data/kings.csv")

kings2 <- read_csv("data/kings.csv")

kings3 <- read.csv2("data/kings.csv")

kings4 <- read_csv2("data/kings.csv")

head(kings1)
head(kings2)
head(kings3)
head(kings4)
```
Answer: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object
 = The functions "read_csv" and "read_csv2" are tidyverse because they produce tibbles. 
2. What is the result of running `class()` on the `kings` object created with a tidyverse function.
 = The object is a tibble, which is a tidyverse version of a data frame. 
3. How many columns does the object have when created with these different functions?
 = The object has 11 columns. We check this by using "ncol(kings)".
4. Show the dataset so that we can see how R interprets each column
 = We show the dataset by using "head()"
```{r eval = TRUE}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE

kings <- read.csv("data/kings.csv",na=c("NULL",""))
ncol(kings)
class(kings)


glimpse(kings)


head(kings)


```


## Calculate the duration of reign for all the kings in your table
Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse.
```{r}
# OUR SOLUTION
kings <- kings %>%
  mutate(duration = End_year- Start_year)

```


## Calculate the average duration of reign for all rulers
```{r}
# OUR SOLUTION

kings %>% 
  select(duration) %>% 
  filter(!is.na(duration)) %>% 
  summarize(meanduration = mean(duration))

```

## How many and which kings enjoyed a longer-than-average duration of reign?

```{r}
#   OUR SOLUTION
kings %>% 
  select(Name, duration) %>% 
  filter(duration > 19.57692) %>% 
    count(duration > 19.57692)
```
25 kings enjoyed a longer-than-average duration of reign. 

## How many days did the three longest-ruling monarchs rule? 
```{r}
# OUR CODE
kings %>%
  arrange(desc(duration)) %>%
  select(Name, duration) %>%
  head(3)
```
The three longest ruling monarchs ruled for 60 years, 51 years and 43 years.
